---
title: "MA5810-Assessment 2"
output: pdf_document
date: "2023-08-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

#### Part A: Exploratory Data Analysis

```{r}
# Load required packages
library(dplyr)
library(caTools)
library(MASS)

# Step 1: Import the data into your session
data_name <- "wdbc.data"
data <- read.csv(data_name, header = FALSE)
head(data)

# Assign column names to the data
col_names <- c("ID", "Diagnosis", paste0("Feature_", 1:30))
colnames(data) <- col_names

# Convert Diagnosis to a binary variable (1 for malignant, 0 for benign)
data$Diagnosis <- as.factor(ifelse(data$Diagnosis == "M", 1, 0))
head(data)
```

#### Part B: Investigate the correlation between the numeric features

```{r}
library(corrplot)
# Task 3: Investigate the correlation between the 30 numeric features
correlation_matrix <- cor(data[, 3:32])
corrplot(correlation_matrix)
```

Inference: There are some highly correlated features in the dataset. Therefore, multicollinearity exists.

#### Part C: Derive secondary feature set using PCA analysis

```{r}
library(caret)
library("factoextra")

# Subset the numeric features only (excluding ID and Diagnosis columns)
numeric_features <- data[, -c(1, 2)]
pca_eval <- prcomp(numeric_features, scale = TRUE)
summary(pca_eval)
 
# Std. Dev of PCA analysis
pca_eval$sdev
 
# Variance is sdev^2
pca_eval.var <- pca_eval$sdev ^ 2
 
# For Scree plot, find proportion of var
varprop <- pca_eval.var / sum(pca_eval.var)

# Scree plot for varprop
plot(varprop, xlab = "Principle Component",
            ylab = "Proportion of Variance",
            ylim = c(0, 1), type = "b",
            main = "Variance Proportion")

# Screr plot of cumulative proportion
plot(cumsum(varprop),
    xlab = "Principle Component",
    ylab = "Cumulative Proportion",
    ylim = c(0, 1), type = "b")
 
which(cumsum(varprop) > 0.9)[1]
```

#### Part D: Form the transformed data

```{r}
# Get the first 6 principal components
pca_components <- pca_eval$x[, 1:4]

# Create a new data frame with the transformed data
transformed_data <- as.data.frame(pca_components)

# Add the ID and Diagnosis columns back to the transformed data
transformed_data <- cbind(data[, c(1, 2)], transformed_data)

head(transformed_data)

```

#### Part E: Fit a Logistic Regression Model on the Transformed data

```{r}
set.seed(42) # For reproducibility
split_new <- sample.split(transformed_data, SplitRatio = 0.9)
train_transformed_data <- subset(transformed_data, split_new == TRUE)
test_transformed_data <- subset(transformed_data, split_new == FALSE)

transformed_model <- glm(Diagnosis ~ ., data = train_transformed_data[, -1], family = binomial)
summary(transformed_model)
```

#### Part F: Prediction and Accuracy of the the model

```{r}

# Predict the diagnoses on the test data using the logistic regression model
logreg_preds <- predict(transformed_model, newdata = test_transformed_data[, -1], type = "response")

# Convert the predicted probabilities to diagnoses
logreg_preds  <- ifelse(logreg_preds > 0.5, 1, 0)

# Compute the accuracy of the logistic regression model on the test data
logreg_accuracy <- mean(logreg_preds == test_transformed_data$Diagnosis)
# Print the accuracy of the logistic regression model
cat("Transformed Model Logistic Regression Accuracy:", logreg_accuracy, "\n")
```
